Learning Reusable Components for Neurally--Guided Bayesian Program Learning

### 2018-05-13 nips_2018.tex
we can do sygus

We propose \emph{learning} the DSL by inducing a library of domain--specific
subroutines. We consider the setting where we have a collection of
related programming tasks, each specified by a set of input/output
examples. %% We do not assume that the tasks are annotated with
%% ground-truth programs.
Starting from a weaker or more
general library of primitives, we give an algorithm for constructing a richer, more powerful,
and better-tuned DSL.
Our algorithm is called \textbf{Explore/Compress/Compile} (\systemEnding)
because it iterates between three different phases: an
\textbf{Explore} phase uses the DSL to explore the space of programs,
searching for ones that solve the tasks; a \textbf{Compress} phase
modifies the DSL by discovering regularities in the programs found by
the previous Explore phase; and a \textbf{Compile} phase, which
improves the program search procedure by training a neural network to
write programs in the current DSL, in the spirit of ``amortized'' or
``compiled'' inference~\cite{le2016inference}.
We call the neural net a \textbf{recognition model} (c.f. Hinton 1995~\cite{hinton1995wake}).
The learned DSL
distills commonalities across programs that solve tasks, helping
the agent solve related program induction problems. The neural
recognition model ensures that searching for programs remains tractable
even as the DSL (and hence the search space for programs) expands.

@@

We apply \system to three domains:
 symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing; and Lisp-style functions on lists.
 For each of these we provide a generic
 set of programming primitives in a Lisp-like language,
 including conditionals, variables, and higher-order functions.
 Our algorithm then discovers
 its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
\begin{wrapfigure}{r}{0.6\textwidth}
    \begin{tabular}{ll}
   \toprule
   Domain&Part of the learned DSL\\\midrule
   %% Circuits&\code{(nand (nand a a) (nand b b))}\\
   %% &\hspace{0.5cm} \emph{(an OR gate)}\\
   Regression& \code{(+ (* real x) real)} \\
      &\hspace{0.5cm} \emph{(a linear function of x)}\\
   Strings& \code{(map (lambda (x) (if (= x a) b x)))}\\
   &\hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
   Lists& \code{(map (lambda (x) (+ x k)) l)}\\
   &\hspace{0.5cm} \emph{(add k to every element of list l)}
   \\\bottomrule
    \end{tabular}
    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
 \end{wrapfigure}

@@

 Our goal is to induce a DSL while finding programs solving each of the
tasks.  We take inspiration primarily from the
Exploration-Compression algorithm for bootstrap
learning~\cite{Dechter:2013:BLV:2540128.2540316}.  Exploration-Compression
alternates between exploring the space of solutions to a set of tasks,
and compressing those solutions to suggest new search primitives
for the next exploration stage.
We extend these ideas into an inference strategy that iterates
through three steps: an \textbf{Explore} cycle uses the current DSL and
recognition model to search for programs that solve the tasks.  The
\textbf{Compress} and \textbf{Compile} cycles update the DSL and the
recognition model, respectively.  Crucially, these steps bootstrap 
off each other (Fig.~\ref{feeding}):

\\\noindent \textbf{Exploration: Searching for programs.}  Our program search
is informed by both the DSL and the recognition model. When these
improve, we find more programs solving the tasks.
\\\noindent\textbf{Compression: Improving the DSL.} We induce the DSL
from the programs found in the exploration phase, aiming
to maximally compress (or, raise the prior probability of) these
programs.  As we solve more tasks, we hone in on richer DSLs that
more closely match the domain.
\\\noindent \textbf{Compilation: Learning a neural recognition model.}  We
update the recognition model by training on two data sources: samples
from the DSL (as in the Helmholtz Machine's ``sleep'' phase), and
programs found by the search procedure during exploration. As the DSL
improves and as search finds more programs, the recognition model gets
both more data to train on, and better data.
\begin{wrapfigure}{r}{0.6\textwidth}
  \begin{tikzpicture}
    \begin{scope}[shift = {(1,-1)}]
    \node[align = center](synthesis) at (6,4) {Search for \\programs};
    \node[align = center](DSL) at (3,1) {DSL};
    \node[align = center](recognitionModel) at (9,1) {Recognition \\model};

@@

    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Exploration)}} (synthesis.west);
    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Exploration)}} (synthesis.east);
  \end{scope}
    \end{tikzpicture}
  \caption{\system solves for programs, the DSL, and a neural network called the \emph{recognition model}. Each of these steps bootstrap off of the others in an iterative inference algorithm.}  \label{feeding}
\end{wrapfigure}

Sec.~\ref{mathematicalFraming} frames this 3-step procedure as probabilistic inference.
Sec.~\ref{explorationSection} explains how we search for programs that solve the tasks;
Sec.~\ref{recognitionSection} explains how we train a neural network to  search for programs; and
Sec.~\ref{grammarInductionSection} explains how we induce a DSL from programs.

\subsection{Probabilistic Framing}\label{mathematicalFraming}

@@

Searching for programs is hard because
of the large combinatorial search space. We ease this difficulty by training a neural recognition model, $q(\cdot |\cdot )$,
during the compilation phase: $q$ is trained to approximate the
posterior over programs, $q(p|x)\propto\probability[x,p|\mathcal{D},\theta}]$,
  thus amortizing the cost of finding programs with high posterior probability.

@@

Now our goal is to search for programs that solve the tasks.  In this
work we use the simple search strategy of enumerating programs from
the DSL  in decreasing order of their probability,
and then checking if an enumerated program $p$ assigns positive
probability to a task ($\probability[x|p] > 0$); if so, we incorporate $p$ 
into the frontier $\mathcal{F}_x$.

@@

The purpose of the recognition model is to accelerate the search for
programs.  It does this by learning to predict programs which are
probable under $(\mathcal{D},\theta)$ while also assigning high
likelihood for a task according to $\probability[x|p]$.  Concretely,
the recognition model $q$ is a neural network that predicts, for each
task $x\in X$, a weight vector $q(x) = \theta^{(x)}\in
\mathbb{R}^{|\mathcal{D}| + 1}$.  Together with the DSL, this defines
a distribution over programs, $\probability[p|\mathcal{D},\theta =
  q(x)]$.  We abbreviate this distribution as $q(p|x)$.  The crucial
aspect of this framing is that the neural network leverages the
structure of the learned DSL, so it is \emph{not} responsible for
generating programs wholesale.  We share this aspect with
DeepCoder~\cite{balog2016deepcoder} and~\cite{menon2013machine}.

We want a recognition model that closely approximates the true posteriors over programs. We formulate this as minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|q(p|x) \right) \right]$,
or equivalently maximizing $  \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]$,
%% \begin{equation*}
%%   \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
%% \end{equation*}
where the expectation is taken over tasks. One could take this expectation
over the observed empirical distribution of tasks,
like how an autoencoder is trained~\cite{hinton2006reducing}; or, one could take this expectation over samples from the generative model, like how a Helmholtz machine is trained~\cite{dayan1995helmholtz}.
We found it useful to maximize both an autoencoder-style objective $\mathcal{L}_{\text{AE}}$ and a Helmholtz-style objective $\mathcal{L}_{\text{HM}}$, giving the  objective for a recognition model, $\mathcal{L}_{\text{RM}} = \mathcal{L}_\text{AE} + \mathcal{L}_\text{HM}$:
\begin{align}\nonumber
\mathcal{L}_{\text{HM}} = \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right]\quad
\mathcal{L}_{\text{AE}} = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
  \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\log q(p|x)\right]
\end{align}
The $\mathcal{L}_{\text{HM}}$ objective is essential for data efficiency:
all of our experiments train \system on only a few hundred tasks, which is too little for
a high-capacity neural network $q$.
Once we bootstrap a $(\mathcal{D},\theta)$,
we can draw unlimited samples from $(\mathcal{D},\theta)$
and train $q$ on those samples.

@@

        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Compress)}}
        %% \STATE Define $Q_x(z) \propto \begin{cases}
        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
        %%   0&x\not \in \mathcal{F}_x
        %% \end{cases}
        \ENDFOR
        \STATE \textbf{return} $\mathcal{D},\theta,q$
      \end{algorithmic}
    \end{algorithm}
  \end{minipage}
\vspace{-0.7cm}\end{wrapfigure}To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
Although this problem may seem 
very similar to estimating the parameters of a probabilistic context free grammar (PCFG),
for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
our DSLs are context-sensitive due to the presence of variables
in the programs and also due to the polymorphic typing system.
In the Supplement we derive a tractable MAP estimator for $\theta$.

\begin{figure*}
\centering  \begin{tabular}{ll}
    \toprule
    Example programs in frontiers&Proposed subexpression\\\midrule
    \begin{tabular}{l}
      \code{(lambda (a b) (fold b (cons "," a)}\\
      \phantom{\code{(lambda }}\code{(lambda (x z) (cons x z))))}\\
      \code{(lambda (a b) (fold a b (lambda (x z) (cons x z))))}
    \end{tabular}&
    \begin{tabular}{l}
      \code{(fold a b (lambda (x z) }\\
      \phantom{\code{(fold a b }}\code{(cons x z)))}
      \end{tabular}
\\\bottomrule
  \end{tabular}
  \caption{The DSL induction algorithm proposes subexpressions of programs to add to the DSL.
    These subexpressions are taken from programs in the frontiers (left column), and can introduce new variables (right column: \code{a} and \code{b}). Here, the proposed subexpression appends two lists.} 
  \label{fragmentExample}
\end{figure*}
Alg.~\ref{mainAlgorithm} describes how we combine program search,
recognition model training, and DSL induction.
\begin{comment}
\subsection{Implementing \system}

@@

\section{Experiments}
%% We first present our experimental results qualitatively,
%% illustrating the learned DSLs.
%% We then show quantitative results (Sec.~\ref{quantitative})
%% comparing \system with lesioned versions
%% of our model that are analogous to prior work.

\subsection{Symbolic Regression}\label{regressionSection}
\begin{wrapfigure}{r}{0.3\textwidth} \newcommand{\functionSize}{0.75cm}\vspace{-0.5cm}
  \includegraphics[width = \functionSize]{figures/functions/0.png}
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/207.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/317.png}\\  
  \includegraphics[width = \functionSize]{figures/functions/432.png}
  \includegraphics[width = \functionSize]{figures/functions/415.png}
  \includegraphics[width = \functionSize]{figures/functions/426.png}
  \includegraphics[width = \functionSize]{figures/functions/597.png}
  \includegraphics[width = \functionSize]{figures/functions/591.png}
  \caption{Recognition model input for symbolic regression. While the DSL learns subroutines for rational functions \& polynomials, the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines is appropriate for explaining the observation.}\label{functions}
\end{wrapfigure}
As a simple first example of how our algorithm works, we apply \system
to symbolic regression problems.  For each of these problems, the
agent observes points along the curve of a function, and must write a
program that fits those points.  We initially equip our learner with
addition, multiplication, and division, and task it with solving
200 symbolic regression problems, each either a polynomial of degree
1--4 or a rational function.  Our recognition model here is a
convolutional network that observes an image of the target function's
graph (Fig.~\ref{functions}) -- visually, different kinds of
polynomials and rational functions produce different kinds of graphs,
and so the recognition model can learn to look at a graph and predict
what kind of function best explains it.  A key difficulty, however, is
that these problems are best solved with programs containing real
numbers.  Our solution to this difficulty is to allow the system to
write programs with real-valued parameters, and then fit those
parameters by automatically differentiating through the programs the
system writes and use gradient descent to fit the parameters.
We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
%% \begin{align*}
%%   \log \probability[x|p] \triangleq& \sum_{n\leq N}\log
%%   \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
%% \end{align*}

@@

\system learns a DSL containing templates for polynomials of different orders, as well as

which lets the algorithm quickly find the functions that are most appropriate to this domain (Tbl.~\ref{regressionDSL}).
We also found that the algorithm discovers programs that minimize the number of continuous degrees of freedom.
For example, it represents the linear function $-x+2$ with the program
\code{(* (+ x }$\mathcal{R}$\code{) }$\mathcal{R}$\code{)} which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}
which has five continuous parameters.
This phenomenon arises from our Bayesian framing -- both the bias towards shorter programs and the likelihood model's BIC penalty.
\begin{wrapfigure}{l}{0.5\textwidth}\centering
\begin{tabular}{l}
  \toprule
  $f_0\,=\,$\code{(+ $\mathcal{R}$)}\\
  $f_1($\code{x}$)\,=\,$\code{(* (* (+ x $\mathcal{R}$) $\mathcal{R}$) x)} \\
  $f_2($\code{x}$)\,=\,$\code{(+ $\mathcal{R}$ ($f_1$ x))} \\
  $f_3($\code{x}$)\,=\,$\code{(* x ($f_2$ x))}\\
  $f_4($\code{x}$)\,=\,$\code{($f_0$ ($f_3$ x))}\\
  $f_5($\code{x}$)\,=\,$\code{(* x ($f_4$ x))}\\
  $f_6($\code{x}$)\,=\,$\code{($f_0$ ($f_4$ x))} (\emph{4th order polynomial})
  \\\bottomrule
\end{tabular}
\caption{Some learned DSL primitives for symbolic regression. The system starts with addition, multiplication, and real numbers, and learns to build polynomials up to 4th order.}\label{regressionDSL}
\end{wrapfigure}

@@

\subsection{Programs that manipulate sequences}\label{sequences}
We apply \system to text editing (Section~\ref{textSection}) and list processing (Section~\ref{listSection}).
For both these domains we initially provide the system with a generic
minimal set of programming primitives,
and use a bidirectional GRU~\cite{cho2014learning} for the recognition model.
The initial set of primitives includes routines commonly found in Lisp or Scheme interpreters: \code{fold}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.

\subsubsection{String Editing}\label{textSection}
Synthesizing programs that manipulate strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
However, this prior work presumes a ready-made DSL,
expertly crafted to suit string editing.
We show \system can instead start out with generic Lisp primitives
and recover many of the higher-level building blocks that have made these
other system successful.
We automatically generated  600 string editing tasks  with 4 input/output examples each (Fig.~\ref{exampleTextProblem}) and model strings as lists of characters.
At first, \system cannot find any correct programs for most of the tasks.
It assembles a DSL (Tbl.~\ref{textPrimitives}) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.

\begin{table}[t]
  \centering
    \begin{tabular}{l}
      \toprule
      $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b (lambda (x y) (cons x y))))))}\\
      \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (a x) (cdr (if (= c x) s a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
      $f_2($\code{s}$) \,=\, $\code{(unfold s empty? car (lambda (z) (}$f_1$\code{ z SPACE)))}\\
        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words})
  \\\bottomrule
    \end{tabular}
  \caption{Some learned DSL primitives for string editing}\label{textPrimitives}
  \end{table}

>>>>

We propose \emph{learning} the DSL by inducing a library of domain--specific
subroutines. We consider the setting where we have a collection of
related programming tasks, each specified by a set of input/output
examples. %% We do not assume that the tasks are annotated with
%% ground-truth programs.
Starting from a weaker or more
general library of primitives, we give an algorithm for constructing a richer, more powerful,
and better-tuned DSL.
Our algorithm is called \textbf{Explore/Compress/Compile} (\systemEnding),
and iterates between three different steps: an
\textbf{Explore} step uses the DSL to explore the space of programs,
searching for ones that solve the tasks; a \textbf{Compress} step
modifies the DSL by discovering regularities in the programs found by
the previous Explore step; and a \textbf{Compile} step, which
improves the program search procedure by training a neural network to
write programs in the current DSL, in the spirit of ``amortized'' or
``compiled'' inference~\cite{le2016inference}.
We call the neural net a \textbf{recognition model} (c.f. Hinton 1995~\cite{hinton1995wake}).
The learned DSL
distills commonalities across programs that solve tasks, helping
the agent solve related program induction problems. The neural
recognition model ensures that searching for programs remains tractable
even as the DSL (and hence the search space for programs) expands.

@@

We apply \system to three domains:
 symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing; and Lisp-style functions on lists.
 For each of these we initially provide a generic
 set of programming primitives in a Lisp-like language.
% including conditionals, variables, and higher-order functions.
 Our algorithm then discovers
 its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
\begin{wrapfigure}{r}{0.6\textwidth}
    \begin{tabular}{ll}
   \toprule
   Domain&Part of the learned DSL\\\midrule
   %% Circuits&\code{(nand (nand a a) (nand b b))}\\
   %% &\hspace{0.5cm} \emph{(an OR gate)}\\
   Regression& \code{(+ (* real x) real)} \\
      &\hspace{0.5cm} \emph{(a linear function of x)}\\
   Strings& \code{(map (lambda (x) (if (= x a) b x)))}\\
   &\hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
   Lists& \code{(map (lambda (x) (+ x k)) l)}\\
   &\hspace{0.5cm} \emph{(add k to every element of list l)}
   \\\bottomrule
    \end{tabular}
    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}\vspace{-1cm}
 \end{wrapfigure}

@@

 Our goal is to induce a DSL while finding programs solving each of the
tasks.  We take inspiration primarily from the
Exploration-Compression algorithm for bootstrap
learning~\cite{Dechter:2013:BLV:2540128.2540316}.  Exploration-Compression
alternates between exploring the space of solutions to a set of tasks,
and compressing those solutions to suggest new search primitives
for the next exploration stage.
We extend these ideas into an inference strategy that iterates
through three steps: an \textbf{Explore} step uses the current DSL and
recognition model to search for programs that solve the tasks.  The
\textbf{Compress} and \textbf{Compile} steps update the DSL and the
recognition model, respectively.  Crucially, these steps bootstrap 
off each other (Fig.~\ref{feeding}):

\begin{wrapfigure}{r}{0.5\textwidth}\vspace{-0.5}
  \begin{tikzpicture}
    \begin{scope}[shift = {(1,-1)}]
    \node[align = center](synthesis) at (6,4) {Search for \\programs};
    \node[align = center](DSL) at (3,1) {DSL};
    \node[align = center](recognitionModel) at (9,1) {Recognition \\model};

@@

    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Exploration)}} (synthesis.west);
    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Exploration)}} (synthesis.east);
  \end{scope}
    \end{tikzpicture}
  \caption{\system solves for programs, the DSL, and a neural network (recognition model). Each of these steps iteratively bootstraps off of the others.}  \label{feeding}\vspace{-0.4cm}
\end{wrapfigure}%
\\\noindent \textbf{Exploration: Searching for programs.}  Our program search
is informed by both the DSL and the recognition model. When these
improve, we find more programs solving the tasks.
\\\noindent\textbf{Compression: Improving the DSL.} We induce the DSL
from the programs found in the exploration phase, aiming
to maximally compress (or, raise the prior probability of) these
programs.  As we solve more tasks, we hone in on DSLs that
more closely match the domain.
\\\noindent \textbf{Compilation: Learning a neural recognition model.}  We
update the recognition model by training on two data sources: samples
from the DSL (as in the Helmholtz Machine's ``sleep'' phase), and
programs found by the search procedure during exploration. As the DSL
improves and as search finds more programs, the recognition model gets
more data to train on, and better data.

\begin{comment}
Sec.~\ref{mathematicalFraming} frames this 3-step procedure as probabilistic inference.
Sec.~\ref{explorationSection} explains how we search for programs that solve the tasks;
Sec.~\ref{recognitionSection} explains how we train a neural network to  search for programs; and
Sec.~\ref{grammarInductionSection} explains how we induce a DSL from programs.
  \end{comment}
\subsection{Hierarchical Bayesian Framing}\label{mathematicalFraming}

@@

Searching for programs is hard because
of the large combinatorial search space. We ease this difficulty by training a neural recognition model, $q(\cdot |\cdot )$,
during the compilation phase: $q$ is trained to approximate the
posterior over programs, $q(p|x)\propto\probability[p|x,\mathcal{D},\theta}]=\probability[x|p]\probability[p|\mathcal{D},\theta}]$,
  thus amortizing the cost of finding programs with high posterior probability.

@@

Now our goal is to search for programs solving the tasks.  We use the simple search strategy of enumerating programs from
the DSL  in decreasing order of their probability,
and then checking if an enumerated program $p$ assigns positive
probability to a task ($\probability[x|p] > 0$); if so, we incorporate $p$ 
into the frontier $\mathcal{F}_x$.

@@

The purpose of the recognition model is to amortize the cost of searching for
programs.  It does this by learning to predict programs which are
probable under $(\mathcal{D},\theta)$ while also assigning high
likelihood for a task according to $\probability[x|p]$.  Concretely,
the recognition model $q$ is a neural network that predicts, for each
task $x\in X$, a weight vector $q(x) = \theta^{(x)}\in
\mathbb{R}^{|\mathcal{D}| + 1}$.  Together with the DSL, this defines
a distribution over programs, $\probability[p|\mathcal{D},\theta =
  q(x)]$.  We abbreviate this distribution as $q(p|x)$.  The crucial
aspect of this framing is that the neural network leverages the
structure of the learned DSL, so it is \emph{not} responsible for
generating programs wholesale.  We share this aspect with
DeepCoder~\cite{balog2016deepcoder} and~\cite{menon2013machine}.

We want a recognition model that closely approximates the true posteriors over programs. We formulate this as minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|q(p|x) \right) \right]$,
or equivalently maximizing $  \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]$,
%% \begin{equation*}
%%   \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
%% \end{equation*}
where the expectation is taken over tasks. One could take this expectation
over the empirical distribution of tasks,
like how an autoencoder is trained~\cite{hinton2006reducing}; or, one could take this expectation over samples from the generative model, like how a Helmholtz machine is trained~\cite{dayan1995helmholtz}.
We found it useful to maximize both an autoencoder-style objective $\mathcal{L}_{\text{AE}}$ and a Helmholtz-style objective $\mathcal{L}_{\text{HM}}$, giving the  objective for a recognition model, $\mathcal{L}_{\text{RM}} = \mathcal{L}_\text{AE} + \mathcal{L}_\text{HM}$:
\begin{align}\nonumber
\mathcal{L}_{\text{HM}} = \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right]\quad
\mathcal{L}_{\text{AE}} = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
  \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\log q(p|x)\right]
\end{align}
The $\mathcal{L}_{\text{HM}}$ objective is essential for data efficiency:
all of our experiments train \system on only a few hundred tasks, which is too little for
a high-capacity neural network $q$.
Once we bootstrap a $(\mathcal{D},\theta)$,
we can draw unlimited samples from $(\mathcal{D},\theta)$
and train $q$ on those samples.

@@

        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Compress)}}
        %% \STATE Define $Q_x(z) \propto \begin{cases}
        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
        %%   0&x\not \in \mathcal{F}_x
        %% \end{cases}
        \ENDFOR
        \STATE \textbf{return} $\mathcal{D},\theta,q$
      \end{algorithmic}
    \end{algorithm}
  \end{minipage}
\vspace{-0.7cm}\end{wrapfigure}To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
Although this  may seem 
very similar to estimating the parameters of a probabilistic context free grammar,
for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
our DSLs are context-sensitive due to the presence of variables
in the programs and also due to the polymorphic typing system.
In the Supplement we derive a tractable MAP estimator for $\theta$.
Putting it all together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
recognition model training, and DSL induction.
\begin{figure*}
\centering  \begin{tabular}{ll}
    \toprule
    Example programs in frontiers&Proposed subexpression\\\midrule
    \begin{tabular}{l}
      \code{(lambda (a b) (fold b (cons "," a)}\\
      \phantom{\code{(lambda }}\code{(lambda (x z) (cons x z))))}\\
      \code{(lambda (a b) (fold a b (lambda (x z) (cons x z))))}
    \end{tabular}&
    \begin{tabular}{l}
      \code{(fold a b (lambda (x z) }\\
      \phantom{\code{(fold a b }}\code{(cons x z)))}
      \end{tabular}
\\\bottomrule
  \end{tabular}
  \caption{The DSL induction algorithm proposes subexpressions of programs to add to the DSL.
    These subexpressions are taken from programs in the frontiers (left column), and can introduce new variables (right column: \code{a} and \code{b}). Here, the proposed subexpression appends two lists.} 
  \label{fragmentExample}
\end{figure*}

\begin{comment}
\subsection{Implementing \system}

@@

%\section{Experiments}
%% We first present our experimental results qualitatively,
%% illustrating the learned DSLs.
%% We then show quantitative results (Sec.~\ref{quantitative})
%% comparing \system with lesioned versions
%% of our model that are analogous to prior work.

\section{Symbolic Regression Experiments}\label{regressionSection}
\begin{wrapfigure}{r}{0.3\textwidth} \newcommand{\functionSize}{0.75cm}\vspace{-0.5cm}
  \includegraphics[width = \functionSize]{figures/functions/0.png}
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/207.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/317.png}\\  
  \includegraphics[width = \functionSize]{figures/functions/432.png}
  \includegraphics[width = \functionSize]{figures/functions/415.png}
  \includegraphics[width = \functionSize]{figures/functions/426.png}
  \includegraphics[width = \functionSize]{figures/functions/597.png}
  \includegraphics[width = \functionSize]{figures/functions/591.png}
  \caption{Recognition model input for symbolic regression. While the DSL learns subroutines for rational functions \& polynomials, the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines is appropriate for explaining the observation.}\label{functions}
\end{wrapfigure}
As a simple first example of how our algorithm works, we apply \system
to symbolic regression problems.  For each of these problems, the
agent observes points along the curve of a function, and must write a
program that fits those points.  We initially equip our learner with
addition, multiplication, and division, and task it with solving
200 symbolic regression problems, each either a polynomial of degree
1--4 or a rational function.  Our recognition model here is a
convolutional network that observes an image of the target function's
graph (Fig.~\ref{functions}) -- visually, different kinds of
polynomials and rational functions produce different kinds of graphs,
and so the recognition model can learn to look at a graph and predict
what kind of function best explains it.  A key difficulty, however, is
that these problems are best solved with programs containing real
numbers.  Our solution to this difficulty is to allow the system to
write programs with real-valued parameters, and then fit those
parameters by automatically differentiating through the programs the
system writes and use gradient descent to fit the parameters.
We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
%% \begin{align*}
%%   \log \probability[x|p] \triangleq& \sum_{n\leq N}\log
%%   \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
%% \end{align*}

@@

\system learns a DSL containing templates for polynomials of different orders, as well as ratios of polynomials (Fig.~\ref{regressionDSL}).
The algorithm also discovers programs that minimize the number of continuous degrees of freedom.
For example, it represents the linear function $-x+2$ with the program
\code{(* real (+ x real))}}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}
which has five continuous parameters.
This phenomenon arises from our Bayesian framing -- both the bias towards shorter programs and the likelihood model's BIC penalty.
\begin{wrapfigure}{l}{0.4\textwidth}\centering
\begin{tabular}{l}
  \toprule
  $f_0($\code{x}$)\,=\,$\code{(* real (+ x real))}\\
  $f_1($\code{x}$)\,=\,$\code{($f_0$ (* x (+ real x))} \\
  \\\bottomrule
\end{tabular}
\caption{Some learned subroutines for symbolic regression. The system starts with addition, multiplication, and real numbers, and learns to build polynomials up to 4th order.}\label{regressionDSL}
\end{wrapfigure}

@@

\section{Sequence manipulating programs}\label{sequences}
We apply \system to text editing (Section~\ref{textSection}) and list processing (Section~\ref{listSection}).
For both these domains we initially provide the system with a generic
minimal set of programming primitives,
and use a bidirectional GRU~\cite{cho2014learning} for the recognition model.
The initial set of primitives includes routines commonly found in Lisp or Scheme interpreters: \code{fold}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.

\subsection{String Editing}\label{textSection}
Synthesizing programs that manipulate strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
However, this prior work presumes a ready-made DSL,
expertly crafted to suit string editing.
We show \system can instead start out with generic Lisp primitives
and recover many of the higher-level building blocks that have made these
other system successful.
We automatically generated  600 string editing tasks  with 4 input/output examples each (Fig.~\ref{exampleTextProblem}) and model strings as lists of characters.
At first, \system cannot find any correct programs for most of the tasks.
It assembles a DSL (Tbl.~\ref{textPrimitives}) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.

How well does the  learned DSL generalized to real text-editing scenarios?
We tested, but did not train, our model on problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
our system solves 32/108 of the problems with an average search time of 11 minutes.
After learning,
our system solves 80/108, and does so much faster,
solving them in an average of 29 seconds.

\begin{table}[t]
  \centering
    \begin{tabular}{l}
      \toprule
      $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b (lambda (x y) (cons x y))))))}\\
      \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (a x) (cdr (if (= c x) s a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
      $f_2($\code{s}$) \,=\, $\code{(unfold s empty? car (lambda (z) (}$f_1$\code{ z SPACE)))}\\
        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words})
  \\\bottomrule
    \end{tabular}
  \caption{Some string editing learned subroutines }\label{textPrimitives}
  \end{table}
